---
title: "Assignment 04"
authors: "Olivia Gomez and Edward Malcolm"
format:
  html:
    self-contained: true
---

```{r, include=FALSE}

library(tidyverse)
library(janitor)
library(ggthemes)
base_url <- "https://www2.census.gov/programs-surveys/demo/datasets/hhp/"
week_url <- "2023/wk52/HPS_Week52_PUF_CSV.zip"

pulse_url <- paste0(base_url, week_url)

dir.create("data")

download.file(
pulse_url,
destfile = "data/pulse52.zip"
)

unzip(zipfile = "data/pulse52.zip", exdir = "data")

pulse <- read_csv("data/pulse2022_puf_52.csv") %>%
  janitor::clean_names()


```

We are using data from the U.S. Census Bureau's Household Pulse Survey - linked [here.](https://www2.census.gov/programs-surveys/demo/datasets/hhp/)

We specifically used data from Week 52.

**Visualization 1**

This visualization explores the income distributions across the four major regions of the U.S.: Northeast, South, Midwest, and West. The four different regions appear to have a normal distribution of income, with the most common income being between \$50,000 and \$74,999. However, it's clear that there are slight differences in distribution, such as the fact that the West skews slightly left towards the higher end of income levels. An example of the applicability of this visualization is that it could inform policies that wish to target different regions to reduce wage disparities.

```{r}

pulse$income <- factor(pulse$income)

pulse %>%
  filter(income != -99, income != -88) %>%
  ggplot() +
  geom_bar(mapping = aes(x = income, alpha = .8), fill = "blue", show.legend = FALSE) +
  facet_wrap(~ region, labeller = labeller(region = c(
    `1` = "Northeast",
    `2` = "South",
    `3` = "Midwest",
    `4` = "West"
  ))) +
  scale_x_discrete(labels = c(
    `1` = "<25K",
    `2` = "25K - 34,999K",
    `3` = "35K - 49,999K",
    `4` = "50K - 74,999K",
    `5` = "75K - 99,999K",
    `6` = "100K - 149,999K",
    `7` = "150K - 199,999K",
    `8` = ">200K"
  )) +
   labs(title = "How Does Income Vary Across the Country?", y = "Count" , x = "Income Bracket ($)", subtitle = "Income Distribution in the Four Regions of the U.S.", caption = "Source: U.S. Census Bureau Household Pulse Survey") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5), panel.grid = element_blank())
```

**Visualization 2**

This visualization plots the average amount spent on groceries for each household size, further disaggregated by region, to illustrate where people are spending more on groceries and the trend between household size and groceries expenditures. As expected, the trend was positive - a larger household spent more on groceries. The interesting spikes in the Northeast and Midwest could flag that further investigation is needed to determine if groceries are significantly more expensive in those regions. This graph could inform nutrition support policies to target those who need it the most.

```{r}

pulse$region <- factor(pulse$region)

pulse %>%
  group_by(thhld_numper, region) %>%
  mutate(avg_spend = mean(tspndfood, na.rm = TRUE)) %>%
  arrange(region, thhld_numper) %>%
  select(thhld_numper, avg_spend, region) %>%
  ggplot(mapping=aes(x = thhld_numper, y = avg_spend, color = region, group = region)) +
  geom_line(linetype = "dashed") + 
  geom_point() +
  geom_text(aes(label = ""), vjust = -0.5) +
  scale_color_manual(values = c("1" = "red", "2" = "blue", "3" = "green", "4" = "purple"),
                     labels = c("1" = "Northeast", "2" = "South", "3" = "Midwest", "4" = "West")) +
  theme_minimal() +
    labs(title = "Does Spending on Groceries Substantially Vary by Region and Household Size?", y = "Average Amount Spent on Groceries ($)", x = "Total Number of People in Household", color = "Region", subtitle = "Average Amount Spent on Groceries Compared to Household Size, by Region", caption = "Source: U.S. Census Bureau Household Pulse Survey") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5), text = element_text(size = 9))

  
```

**Visualization 3**

This visualization shows the amount of expected foreclosures over the next two months along with their assessed likelihood. The visualizations are further distributed by race, specifically white, black, Asian, and other. The data shows that the majority of expected foreclosures, at all levels of likelihood, are among those who identify as white. The next step researchers would want to take from this visualization is to account for the share of the housing market each race takes up, thereby being able to determine which race is most likely to experience foreclosures.

```{r}
pulse %>%
  filter(forclose != -99, forclose != -88) %>%
  ggplot(aes(x = forclose, fill = factor(rrace))) +  
  geom_bar() +
  labs(title = "Does Foreclosure Likelihood Vary By Race?", y = "Count", x = "Response", fill = "Race", subtitle = "Likelihood of Foreclosure on Home in the Next Two Months, by Race", caption = "Source: U.S. Census Bureau Household Pulse Survey") +
scale_x_discrete(
    limits = c("1", "2", "3", "4"),
    labels = c(
      "1" = "Very Likely",
      "2" = "Somewhat Likely",
      "3" = "Not Very Likely",
      "4" = "Not at All Likely"
    )) +
  scale_fill_manual(values = c("1" = "red", "2" = "blue", "3" = "green", "4" = "purple"),
                     labels = c("1" = "White", "2" = "Black", "3" = "Asian", "4" = "Any other Race")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5))


```

**Visualization 4**

This visualization shows the average number of children respondents had based on their education level. Educational levels included: less than high school, some high school, high school degree, some college, associate’s degree, bachelor’s degree, and graduate degree. Perhaps as expected, respondents with less education had a notably higher average number of children (1.12), with that rate drastically falling and leveling out between 0.6 and 0.7 with a high school degree or higher. This insight can shape which demographics should receive varying types of resources.

```{r}
pulse %>%
  group_by(eeduc) %>%
  summarize(avg_num_child = mean(thhld_numkid, na.rm = TRUE)) %>%
  ggplot(mapping = aes(x = factor(eeduc), y = avg_num_child)) +
  geom_col(fill = "green") +
  geom_text(aes(label = round(avg_num_child, 2)), vjust = -0.5, size = 3) +
  labs(title = "Does Education Level Impact the Number of Children A Respondent Has?", y = "Average Number of Children" , x = "Education Level", subtitle = "Average Number of Children by Educational Attainment", caption = "Source: U.S. Census Bureau Household Pulse Survey") +
  scale_x_discrete(labels = c(
    `1` = "Less than high school",
    `2` = "Some high school",
    `3` = "High school Degree",
    `4` = "Some college",
    `5` = "Associate’s degree",
    `6` = "Bachelor's degree",
    `7` = "Graduate degree"
  )) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5), text = element_text(size = 9))

```
